Debugging by \textit{lastChange}

### 2010-10-20 papers/pldi2011/sigplanconf-template.tex

%---------------------------------------------------------------------------------------------------
\section{Introductory example}
We illustrate the \textit{lastChange} functionality by a simple example. The example demonstrates a buggy JavaScript code in a html page (Figure ~\ref{fig:js-code}). The page contains a button (line 40) which shows the value of \texttt{myObject.myProperty}. It is supposed that whenever user clicks on the button, the value of \texttt{myObject.myProperty} and therefore button's label is increased by one. Once the page is loaded for the first time the button shows \texttt{1} as the initial value of \texttt{myObject.myProperty}. The bug happens when user clicks on the button and instead of \texttt{2}, \texttt{0} appears.

Once user clicks on the button, \texttt{onClick} function (line 13) is called. This function increases the value of \texttt{myObject.myProperty} by one (line 15) and calls \texttt{updateButton} function which updates button's text to the new value (line 22). To start debugging, the developer sets a breakpoint on line 22 and once the button is clicked execution is paused at line 22. Figure ~\ref{fig:example1} shows the Firebug debugger while the execution is paused. Firebug has several panels (e.g., html, css, script, dom, etc.) that each one demonstrate one aspect of JavaScript execution. The script panel contains the list of all loaded source files and regular debugging facilities such as setting breakpoints and stepping. The right panel in the script panel shows the program state and developer can examine object and variable values. In our case, the \texttt{myObject.myProperty} value at the paused point is zero which is obviously a wrong value.

\begin{figure}[htp]
\begin{verbatim}
1 <html>
...
5   <script type="text/javascript">
6    myObject = {myProperty : 1};
7    myCondition = {value : 1};
...
13   function onClick(){
14     foo();
15     myObject.myProperty++;
16     bar();
17     ...
18     updateButton();
19   }
20   function updateButton(){
21     var myParagraph =
          document.getElementById("myButton");
22     myButton.innerHTML = myObject.myProperty;
23   }   
24   function foo(){
25  	 myCondition.value = oldValue;
26   }  
27   function bar(){ 
28     if (!myCondition.value)
29         myObject.myProperty = 0;
30   }
31  </script> 
...
40  <button id="myButton" onclick="onClick()">
41  	1 
42  </button>
43 </html>
\end{verbatim}
\caption{The buggy JavaScript code in a html page.}
\label{fig:js-code}
\end{figure}

@@

To apply backward search strategy for locating defects, the developer first needs to know the origin to the wrong value. To achieve this goal using breakpoints, the developer should search code to find and set breakpoint at all possible places that \texttt{myObject.myProperty} might get a new value. Having the fact that an object and property can be accessed and changed through different names and methods, we realize that the difficulty of this task. Even if the developer could sucessfully finish this step (There is no simple way to check it!), the next step is monitoring the re-execution and collecting data at breakpoint hits before reaching line 22. All this work should be repeated if an origin was missed in the first step or if the data was collected in the second step is insufficient.

In contrast, we propose a high-level function in debugger, \textit{lastChange}, which provides the answer without demanding anymore effort from the developer. By right clicking on \texttt{myObject.myProperty} in the right panel, the developer can run \textit{lastChange} command (Figure ~\ref{fig:example2}). Debugger considers the paused point at the breakpoint as the reproduction point, and re-executes the buggy execution, pauses at the reproduction point and shows line 29 in a new panel (Figure ~\ref{fig:example3}). The program state debugger shows on the right panel belongs to the past. Therefore the right panel just shows the partially collected data. If the developer needs some values which are not available in the right panel, debugger re-executes another time and collects and shows the requested data.

@@

\subsection{\textit{lastChange} on object propery}
To simplify the algorithm explanation and avoid technical details, we assume two basic operations and later we explain the details of this two operations. The first opertion is \texttt{objectId()} which gets a JavaScript object and returns an integer as its identifier. This identifier is unique for every object during one execution. The second operation is \texttt{setPropertyChangeHook()} which gets function as the callback and a string as the property name like \texttt{foo}. Whenever property \texttt{foo} in any object changes the callback will be called. The callback function receives a reference to the owner of \texttt{foo}.

@@

Once developer asks for the last change of \texttt{bar.foo} at the reproduction point, debugger calls \texttt{setPropertyChangeHook()} with \texttt{foo} as the property name and re-executes the program. Whenever \texttt{foo} changes and the callback function is called, debugger first calls \texttt{objectId()} on the \texttt{foo} owner object. Then it stores owner id and stack frame locations. The reason we keep object id instead of a reference to the object itself is to let garbage collector to destroy dead objects. Whenever the execution reaches the reproduction point it looks at the history of \texttt{foo} changes and finds the last \texttt{foo} change with the same object id as \texttt{bar} id at the reproduction point. Figure ~\ref{fig:foo-changes1} shows the list of property \texttt{foo} change events in a hypothetical execution. \texttt{bar} id at the reproduction point is 1010, so the last change of \texttt{bar.foo} is the forth column.

\subsubsection{\texttt{objectId()}}
The object id is kept as a property of the object, \texttt{\_objectid}, which is not enumerable. Therefore, it will not appear in \texttt{for(property in object)} loops and has no effect on the executing program. Once \texttt{objectId()} is called on an object and the object has not \texttt{\_objectId} property, this property is defined by calling \texttt{defineProperty} JavaScript function. As a parameter it can be specified that the defined property is not enumerable. Then it is enough to assign a new generated id to this property.

@@

The scope id is kept as a variable in the scope, \texttt{\_scopeId}. As we said, a hook is set on the first instruction of the function defines the variable. Whenever this hook hits,  \texttt{\_scopeId} variable is set by \texttt{eval()} function. For example evaluating \texttt{var \_scopeId = 50} will create a variable with name \texttt{\_scopeId} and value \texttt{50} in the local scope.

@@

\subsection{Tracking object creation}
Now we discuss two mentioned assumptions. In the standard JavaScript every object has a \texttt{watch()} function which receives two parameters, a property name and another function as a hook. Whenever the property with the given name changes the hook function is called. The issue with using this function is that it can only be applied to available objects. However, at the beginning of re-execution there is no object available. To solve this issue we use a feature in Firefox JavaScript engine, setting a flag gives us the place the object created (e.g., myFile.js, line 24). Though this is a limited data it can help us to get a reference to the object once it is created. Therefore, at the reproduction point, debugger gets the creation location of \texttt{bar} and once the execution reaches to that location it gets a reference to the created object and sets a unique id for it.

@@

\section{Evaluation}

>>>>

%---------------------------------------------------------------------------------------------------
\section{Introductory example}

We illustrate the \textit{lastChange} functionality by a simple
example. The example demonstrates a buggy JavaScript code in a HTML
page (Figure ~\ref{fig:js-code}). The page contains a button (line 40)
showing the value of \texttt{myObject.myProperty}.  When the user
clicks on the button, the \texttt{onClick} function (line 13) is
called. This function increases the value of
\texttt{myObject.myProperty} by one (line 15) and calls
\texttt{updateButton} function which updates button's text to the new
value (line 22).  Once the page is loaded for the first time the
button shows \texttt{1} as the initial value of
\texttt{myObject.myProperty}.  In practice when the user clicks on the
button, \texttt{0} appears instead of \texttt{2}: there is a bug.

Two other functions are called in \texttt{onClick()}, \texttt{foo()}
and \texttt{bar()}. As developers we often encounter function calls
which seem perpheral to our current concern; they may have been added
by another developer, or we may have forgotten their exact properties
or those properties may have changed, and so on. The difference
between what we expect these functions to do, e.g. nothing
interesting, and what they do in practice may cause bugs.

To start debugging, the developer sets a breakpoint
on line 22 and once the button is clicked execution is paused at line
22. Figure ~\ref{fig:example1} shows the Firebug debugger while the
execution is paused. Firebug has several panels (e.g., HTML, CSS,
Script, DOM, etc.) that each demonstrate one aspect of the Web page.
The Script panel contains the list of all loaded source
files and regular debugging facilities such as setting breakpoints and
stepping. To the right of the script panel, the Watch panel shows the program state
where the developer can examine object and variable values. In our case, the
\texttt{myObject.myProperty} value at the paused point is zero. We expected this value to be \texttt{2}.

\begin{figure}[htp]
\begin{verbatim}
1 <html>
...
5   <script type="text/javascript">
6    myObject = {myProperty : 1};
7    myCondition = {value : 1};
...
13   function onClick(){
14     foo();
15     myObject.myProperty++;
16     bar();
17     ...
18     updateButton();
19   }
20   function updateButton(){
21     var myParagraph =
          document.getElementById("myButton");
22     myButton.innerHTML = myObject.myProperty;
23   }   
24   function foo(){
25  	 myCondition.value = oldValue;
26   }  
27   function bar(){ 
28     if (!myCondition.value)
29         myObject.myProperty = 0;
30   }
31  </script> 
...
40  <button id="myButton" onclick="onClick()">
41  	1 
42  </button>
43 </html>
\end{verbatim}
\caption{A Web page containing JavaScript code. Some lines not related to our paper have been elided.}
\label{fig:js-code}
\end{figure}

@@

To apply backward search strategy for locating defects, the developer
first needs to know the origin to the wrong value. To achieve this
goal using breakpoints, the developer should search code to find and
set breakpoint at all possible places that
\texttt{myObject.myProperty} might get a new value.  However, an
object and property can be accessed and changed through different
names and methods. There is no simple way to identify these alias or
even their total number.  The developer can make a good guess and set
breakpoints on lines where the property seems to be changed. Then they
re-execute the program and examine the state looking for values that
may lead to the incorrect value observed at line 22. All this work
must be repeated if a new alias is discovered or if the some
information related to the buggy result was missed while stopped on
one of the breakpoints.

In contrast, we propose a high-level function in debugger,
\textit{lastChange}, which provides the answer without tedious manual
effort from the developer. By right clicking on
\texttt{myObject.myProperty} in the Watch panel, the developer can run
\textit{lastChange} command (Figure ~\ref{fig:example2}). Debugger
re-executes the program and halts again at the breakpoint on line 22.
However, it shows a new QP panel, centered on the source at line 29
(Figure ~\ref{fig:example3}), the point of \textit{lastChange}.  To
the right, the TraceData panels shows values of properties from the
program state when it passed through line 29.  These two panels
resemble the Script and Watch panels, but they show data collected by
the debugger at an execution point which has now past: these are
'traces' or 'logs' of information collected during the re-execution.

@@

\subsection{\textit{lastChange} on object propery}
To simplify the algorithm explanation and avoid technical details, we assume two basic operations and later we explain the details of these two operations. The first opertion is \texttt{objectId()} which gets a JavaScript object and returns an integer as its identifier. This identifier is unique for every object during one execution. The second operation is \texttt{setPropertyChangeHook()} which gets function as the callback and a string as the property name like \texttt{foo}. Whenever property \texttt{foo} in any object changes the callback will be called. The callback function receives a reference to the owner of \texttt{foo}.

@@

Once developer asks for the last change of \texttt{bar.foo} at the reproduction point, debugger calls \texttt{setPropertyChangeHook()} with \texttt{foo} as the property name and re-executes the program. Whenever \texttt{foo} changes and the callback function is called, debugger first calls \texttt{objectId()} on the \texttt{foo} owner object. Then it stores owner id and stack frame locations. The reason we keep object id instead of a reference to the object itself is to let garbage collector to destroy dead objects. Whenever the execution reaches the reproduction point it looks at the history of \texttt{foo} changes and finds the last \texttt{foo} change with the same object id as \texttt{bar} id at the reproduction point. Figure ~\ref{fig:foo-changes1} shows the list of property \texttt{foo} change events in a hypothetical execution. \texttt{bar} id at the reproduction point is 1010, so the last change of \texttt{bar.foo} is the fourth column.

@@

@@

\subsection{Algorithm technical details}
\subsubsection{\texttt{objectId()}}
The object id is kept as a property of the object, \texttt{\_objectid}, which is not enumerable. Therefore, it will not appear in \texttt{for(property in object)} loops and has no effect on the executing program. Once \texttt{objectId()} is called on an object and the object has not \texttt{\_objectId} property, this property is defined by calling \texttt{defineProperty} JavaScript function. As a parameter it can be specified that the defined property is not enumerable. Then it is enough to assign a new generated id to this property.

\subsubsection{\texttt{scopeId()}}
The scope id is kept as a variable in the scope, \texttt{\_scopeId}. As we said, a hook is set on the first instruction of the function defines the variable. Whenever this hook hits,  \texttt{\_scopeId} variable is set by \texttt{eval()} function. For example evaluating \texttt{var \_scopeId = 50} will create a variable with name \texttt{\_scopeId} and value \texttt{50} in the local scope.

\subsubsection{Tracking object creation}
Now we discuss two mentioned assumptions. In the standard JavaScript every object has a \texttt{watch()} function which receives two parameters, a property name and another function as a hook. Whenever the property with the given name changes the hook function is called. The issue with using this function is that it can only be applied to available objects. However, at the beginning of re-execution there is no object available. To solve this issue we use a feature in Firefox JavaScript engine, setting a flag gives us the place the object created (e.g., myFile.js, line 24). Though this is a limited data it can help us to get a reference to the object once it is created. Therefore, at the reproduction point, debugger gets the creation location of \texttt{bar} and once the execution reaches to that location it gets a reference to the created object and sets a unique id for it.

@@

\section{Evaluation}
\subsection{Execution overhead analysis}
%If we assume for every change event k additionial instructions should be executed and the number of instructions executed in one execution is n and the number of change %events is m then the total amount of instructions is n+mk. The overhead is mk/n. for m=10^3, k=10^2 and n=10^10 the overhead is very low about 10^-5.
